---
title: "gene modules scores/expression plots"
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: readable
    code_folding: none
---

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(glue)
  library(cowplot)
  library(RColorBrewer)
  library(ggbeeswarm)
})
```

output dir

```{r}
subset_label = "246samples"

num_var_genes = 10000
pca_dims = 10

n_clust = 20

out_dir = glue("./out-modules-scores")
if (!dir.exists(out_dir)) { dir.create(out_dir) }
out_prefix = glue("{out_dir}/{subset_label}.coding{num_var_genes}.modules.dim{pca_dims}.k{n_clust}")

# modules_csv = glue("./out-modules-generate/{subset_label}.coding{num_var_genes}.modules.dim{pca_dims}.k{n_clust}.genes.csv")
modules_csv = "data/rna-genes-modules.csv"
if (!file.exists(modules_csv)) stop()
# scores_csv = glue("./out-modules-generate/{subset_label}.coding{num_var_genes}.modules.dim{pca_dims}.k{n_clust}.scores.csv")
scores_csv = "data/rna-modules-scores-n246.csv"
if (!file.exists(scores_csv)) stop()
```

# import data

import sample info

```{r message=FALSE, warning=FALSE}
samples_tbl = read_csv("data/rna-samples-n246.csv", show_col_types = FALSE) %>%
  rename(sample_name = `#SAMPLE`) %>%
  mutate(progtype = str_c(type, "-", progtype)) %>%
  arrange(sample_name)
```

```{r}
samples_tbl
```

import modules

```{r message=FALSE, warning=FALSE}
modules_tbl = read_csv(modules_csv, show_col_types = FALSE)
```

```{r}
head(modules_tbl)
```

```{r}
modules_tbl %>% count(module)
```

import module scores

```{r}
scores_tbl = read_csv(scores_csv, show_col_types = FALSE)
```

```{r}
head(scores_tbl)
```

# clean up

```{r}
table(samples_tbl$progtype)
```

```{r}
samples_tbl$progtype = str_replace(samples_tbl$progtype, "Recurrence-SecondaryPrimary", "SecondaryPrimary")
table(samples_tbl$progtype)
```

reorder groups

```{r}
progtype_order =
  c(
    "N-NoRecurrence", "N-SecondaryPrimary", "N-Recurrence-Locoregional", "N-Recurrence-Systemic",
    "T-NoRecurrence", "T-SecondaryPrimary", "T-Recurrence-Locoregional", "T-Recurrence-Systemic"
  )
samples_tbl$progtype = factor(samples_tbl$progtype, levels = progtype_order)
levels(samples_tbl$progtype)
```

```{r}
progtype_colors =
  c(
    ggsci::rgb_material("blue", reverse = TRUE, n = 5)[4:1],
    ggsci::rgb_material("red", reverse = TRUE, n = 5)[4:1]
  )
names(progtype_colors) = progtype_order
progtype_colors
```

get average expression per module

```{r}
modules_list = split(x = modules_tbl$gene, f = modules_tbl$module)
str(modules_list)
```

```{r}
scores_mat = scores_tbl %>% column_to_rownames("module") %>% as.matrix()
scores_mat[, 1:5]
```

```{r}
scores_mat %>% quantile() %>% round(3)
```

# plot

combine samples and scores

```{r}
scores_long_tbl =
  scores_mat %>% t() %>% as_tibble(rownames = "sample_name") %>%
  pivot_longer(!sample_name, names_to = "module", values_to = "exp")
head(scores_long_tbl)
```

```{r}
plot_tbl = inner_join(samples_tbl, scores_long_tbl, by = "sample_name")
plot_tbl
```

filter table by followup time

```{r}
table(plot_tbl$progression5y)
```

```{r}
plot_5y_tbl = plot_tbl %>% filter(progression5y != "UnkRecurrence")
table(plot_5y_tbl$progression5y)
```

```{r}
plot_sp5y_tbl = plot_5y_tbl %>% filter(progression5y != "SecondaryPrimary")
table(plot_sp5y_tbl$progression5y)
```

```{r}
n_distinct(plot_tbl$sample_name)
n_distinct(plot_5y_tbl$sample_name)
```

```{r}
color_scheme = c(brewer.pal(9, "Set1"), brewer.pal(8, "Accent"), brewer.pal(8, "Dark2"))
```

```{r}
subtitle_text = glue("{nrow(samples_tbl)} Samples, {nrow(modules_tbl)} Genes")
subtitle_text
```

adjust plot settings based on the number of modules

```{r}
num_modules = n_distinct(plot_tbl$module)
pt_size = 0.3
box_width = 0.3
facet_rows = 2
if (num_modules > 20) {
  pt_size = 0.2
  box_width = 0.5
  facet_rows = 4
}
```

plot tumor-normal

```{r}
vln_plot =
  plot_tbl %>%
  ggplot(aes(x = type, y = exp)) +
  geom_violin(aes(fill = type), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = color_scheme[c(2, 1)]) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.tumor.png"), plot = vln_plot,
          base_height = 6, base_width = 11)
save_plot(glue("{out_prefix}.scores.tumor.pdf"), plot = vln_plot,
          base_height = 6, base_width = 11)
vln_plot
```

plot tumor and recurrence (4 groups)

```{r}
vln_plot =
  plot_tbl %>%
  ggplot(aes(x = type_progression, y = exp)) +
  geom_violin(aes(fill = type_progression), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = color_scheme) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.type-progression.png"), plot = vln_plot,
          base_height = 6, base_width = 12)
save_plot(glue("{out_prefix}.scores.type-progression.pdf"), plot = vln_plot,
          base_height = 6, base_width = 12)
```

plot tumor and recurrence (4 or 6 groups)

```{r}
vln_plot =
  plot_tbl %>%
  ggplot(aes(x = type_progression5y, y = exp)) +
  geom_violin(aes(fill = type_progression5y), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = color_scheme) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.type-progression5y.png"), plot = vln_plot,
          base_height = 6, base_width = 12)
save_plot(glue("{out_prefix}.scores.type-progression5y.pdf"), plot = vln_plot,
          base_height = 6, base_width = 12)
```

plot tumor and recurrence type (5-year subset)

```{r}
vln_plot =
  plot_tbl %>%
  ggplot(aes(x = progtype, y = exp)) +
  geom_violin(aes(fill = progtype), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = progtype_colors) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.progtype.png"), plot = vln_plot,
          base_height = 6, base_width = 12)
save_plot(glue("{out_prefix}.scores.progtype.pdf"), plot = vln_plot,
          base_height = 6, base_width = 12)
```

plots for 5-year subset

```{r}
type_progression_colors = progtype_colors[c(1,4,5,8)]
names(type_progression_colors) = sort(unique(plot_5y_tbl$type_progression5y))
type_progression_colors
```

plot tumor and recurrence (5-year subset)

```{r}
subtitle_text = glue("{n_distinct(plot_sp5y_tbl$sample_name)} Samples, {nrow(modules_tbl)} Genes")

vln_plot =
  plot_sp5y_tbl %>%
  ggplot(aes(x = type_progression5y, y = exp)) +
  geom_violin(aes(fill = type_progression5y), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = type_progression_colors) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.type-progression5y.png"), plot = vln_plot,
          base_height = 6, base_width = 12)
save_plot(glue("{out_prefix}.scores.type-progression5y.pdf"), plot = vln_plot,
          base_height = 6, base_width = 12)
```

```{r}
# subtitle_text = glue("{n_distinct(plot_5y_tbl$sample_name)} of {n_distinct(plot_tbl$sample_name)} Samples, {nrow(modules_tbl)} Genes")
subtitle_text = glue("{n_distinct(plot_5y_tbl$sample_name)} Samples, {nrow(modules_tbl)} Genes")
subtitle_text
```

plot tumor and recurrence type (5-year subset)

```{r}
vln_plot =
  plot_5y_tbl %>%
  ggplot(aes(x = progtype, y = exp)) +
  geom_violin(aes(fill = progtype), scale = "width", color = NA) +
  geom_quasirandom(bandwidth = 0.8, color = "gray10", size = pt_size, alpha = 0.5) +
  geom_boxplot(width = box_width, color = "gray20", fill = "white", outlier.colour = NA, position = position_dodge(0.9)) +
  scale_fill_manual(values = progtype_colors) +
  labs(
    title = "Module Scores",
    subtitle = subtitle_text,
    x = "Group",
    y = "Score"
  ) +
  theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    legend.title = element_blank()
    # panel.grid.major.y = element_line(colour = "gray50", linetype = "dotted")
  ) +
  facet_wrap(vars(module), nrow = facet_rows, scales = "free", strip.position = "bottom")
save_plot(glue("{out_prefix}.scores.progtype.5y.png"), plot = vln_plot,
          base_height = 6, base_width = 12)
save_plot(glue("{out_prefix}.scores.progtype.5y.pdf"), plot = vln_plot,
          base_height = 6, base_width = 12)
```

calculate p-values for all combinations of subtypes

```{r}
group_combinations = combn(unique(plot_tbl$type), m = 2, simplify = TRUE)

stats_tbl = tibble()
for (module_name in rownames(scores_mat)) {
  # scores_mat[module, ] = colMeans(scores_mat[modules_list[[module]], ])
  for (combination_num in 1:ncol(group_combinations)) {
    
    group1 = group_combinations[1, combination_num] %>% as.character()
    group2 = group_combinations[2, combination_num] %>% as.character()
    # message(glue("comparison : {group1} vs {group2}"))
    
    # calculate stats
    val_group1 = plot_tbl %>% filter(module == module_name, type == group1) %>%
      pull(exp)
    val_group2 = plot_tbl %>% filter(module == module_name, type == group2) %>%
      pull(exp)
    p_val = wilcox.test(x = val_group1, y = val_group2)$p.value
    if (is.na(p_val)) { p_val = 1 }
    if (p_val > 0.001) {
      p_val = round(p_val, 3)
    } else {
      p_val = scales::scientific(p_val, digits = 3)
    }
    # message(glue("p : {p_val}"))
    
    stats_tbl = stats_tbl %>%
      bind_rows(
        c(module = module_name,
          min_val = round(min(val_group1, val_group2), 3),
          max_val = round(max(val_group1, val_group2), 3),
          group1 = group1,
          group2 = group2,
          group1_n = length(val_group1),
          group2_n = length(val_group2),
          group1_median = round(median(val_group1), 3),
          group2_median = round(median(val_group2), 3),
          p_val = p_val)
      )
    
  }
}
stats_tbl = stats_tbl %>%
  mutate_at(
    c("min_val", "max_val", "group1_n", "group2_n", "group1_median", "group2_median"),
    as.numeric
  )
write_csv(stats_tbl, glue("{out_prefix}.stats.tumor.csv"))
```

```{r}
group_combinations = combn(unique(plot_5y_tbl$progtype), m = 2, simplify = TRUE)

stats_tbl = tibble()
for (module_name in names(modules_list)) {
  # zscore_avg_mat[module, ] = colMeans(zscore_mat[modules_list[[module]], ])
  for (combination_num in 1:ncol(group_combinations)) {
    
    group1 = group_combinations[1, combination_num] %>% as.character()
    group2 = group_combinations[2, combination_num] %>% as.character()
    # message(glue("comparison : {group1} vs {group2}"))
    
    # calculate stats
    val_group1 = plot_5y_tbl %>% filter(module == module_name, progtype == group1) %>%
      pull(exp)
    val_group2 = plot_5y_tbl %>% filter(module == module_name, progtype == group2) %>%
      pull(exp)
    p_val = wilcox.test(x = val_group1, y = val_group2)$p.value
    if (is.na(p_val)) { p_val = 1 }
    if (p_val > 0.001) {
      p_val = round(p_val, 3)
    } else {
      p_val = scales::scientific(p_val, digits = 3)
    }
    # message(glue("p : {p_val}"))
    
    stats_tbl = stats_tbl %>%
      bind_rows(
        c(module = module_name,
          group1 = group1,
          group2 = group2,
          group1_n = length(val_group1),
          group2_n = length(val_group2),
          group1_median = round(median(val_group1), 3),
          group2_median = round(median(val_group2), 3),
          p_val = p_val)
      )
    
  }
}
write_csv(stats_tbl, glue("{out_prefix}.stats.progtype.5y.csv"))
```

save

```{r}
write_csv(modules_tbl, glue("{out_prefix}.genes.csv"))
write_csv(scores_tbl, glue("{out_prefix}.scores.csv"))
```
